Stepwise Multiple Regressions

### 2021-02-10 Stepwise Multiple Regressions.Rmd
this is a hot mess :(

```{r, echo = FALSE, message = FALSE, warning = FALSE}
#Packages
library(tidyverse)
library(MASS) #Stepwise regression package

@@

@@

aoc_setup_select <- aoc_setup %>% 
  mutate(log10(dose.particles.mL.master)) %>% 
  mutate(log10(dose.mg.L.master)) %>%
  mutate(log10(dose.um3.mL.master)) %>%
  mutate(log10(size.length.um.used.for.conversions)) %>%
  mutate(log10(exposure.duration.d))

@@

# Model: All Independent Variables, Response Variable: Effect (Y/N)

@@

#Select the columns that we want to feed into the model for simplicity
aoc_setup_select_1 <- aoc_setup_select %>% 
  dplyr::select(
    `log10(dose.particles.mL.master)`, 
    `log10(dose.mg.L.master)`,
    `log10(dose.um3.mL.master)`,
    `log10(size.length.um.used.for.conversions)`, 
    size_f,     
    poly_f, 
    shape_f,
    org_f,
    life_f,
    bio_f,
    `log10(exposure.duration.d)`,
    acute.chronic_f,
    charge_f,
    zetapotential.mV,
    particle_source_f,
    effect_f)

#Convert all data to numerical data
aoc_setup_select_num <- data.matrix(aoc_setup_select_1)

#Convert data matrix into a data frame
x <- as.data.frame(aoc_setup_select_num)

#Drop NAs to get model to be happy - this might be sketchy!?
y<- na.omit(x)

#Fit the full model
Full_Model <- lm(effect_f ~., data = y,
                 na.action = na.omit)
summary(Full_Model)
```

@@

```{r, echo = FALSE}
#Stepwise model
step.model <- stepAIC(Full_Model, direction = "both", 
                      trace = FALSE)

@@

## Remove Charge and Zeta Potential

Missing categories (NA) means that some data are incomplete. This is likely charge and zeta potential which are only in the database for PS and PC spheres. Removed from model to retain more data.

@@

#Select the columns that we want to feed into the model for simplicity
aoc_setup_select2 <- aoc_setup_select %>% 
  dplyr::select(
    `log10(dose.particles.mL.master)`, 
    `log10(dose.mg.L.master)`,
    `log10(dose.um3.mL.master)`,
    `log10(size.length.um.used.for.conversions)`, 
    size_f,     
    poly_f, 
    shape_f,
    org_f,
    life_f,
    bio_f,
    `log10(exposure.duration.d)`,
    acute.chronic_f,
    # charge_f,
    # zetapotential.mV,
    particle_source_f,
    effect_f)

#Convert all data to numerical data
aoc_setup_select_num <- data.matrix(aoc_setup_select2)

#Convert data matrix into a data frame
x <- as.data.frame(aoc_setup_select_num)

#Drop NAs to get model to be happy - this might be sketchy!?
y<- na.omit(x)

#Fit the full model
Full_Model <- lm(effect_f ~., data = y,
                 na.action = na.omit)
summary(Full_Model)
```

@@

## Remove Exposure Duration

Exposure duration not significant - removed from model.

#Select the columns that we want to feed into the model for simplicity
aoc_setup_select3 <- aoc_setup_select %>% 
  dplyr::select(
    `log10(dose.particles.mL.master)`, 
    `log10(dose.mg.L.master)`,
    `log10(dose.um3.mL.master)`,
    `log10(size.length.um.used.for.conversions)`, 
    size_f,     
    poly_f, 
    shape_f,
    org_f,
    life_f,
    bio_f,
    # `log10(exposure.duration.d)`,
    acute.chronic_f,
    # charge_f,
    # zetapotential.mV,
    particle_source_f,
    effect_f)

#Convert all data to numerical data
aoc_setup_select_num <- data.matrix(aoc_setup_select3)

#Convert data matrix into a data frame
x <- as.data.frame(aoc_setup_select_num)

#Drop NAs to get model to be happy - this might be sketchy!?
y<- na.omit(x)

#Fit the full model
Full_Model <- lm(effect_f ~., data = y,
                 na.action = na.omit)
summary(Full_Model)
```

```{r, echo = FALSE}
#Stepwise model
step.model <- stepAIC(Full_Model, direction = "both", 
                      trace = FALSE)

## Remove Size (binned)

Size (binned) not significant - removed from model.

#Select the columns that we want to feed into the model for simplicity
aoc_setup_select4 <- aoc_setup_select %>% 
  dplyr::select(
    `log10(dose.particles.mL.master)`, 
    `log10(dose.mg.L.master)`,
    `log10(dose.um3.mL.master)`,
    `log10(size.length.um.used.for.conversions)`, 
    # size_f,     
    poly_f, 
    shape_f,
    org_f,
    life_f,
    bio_f,
    # `log10(exposure.duration.d)`,
    acute.chronic_f,
    # charge_f,
    # zetapotential.mV,
    particle_source_f,
    effect_f)

#Convert all data to numerical data
aoc_setup_select_num <- data.matrix(aoc_setup_select4)

#Convert data matrix into a data frame
x <- as.data.frame(aoc_setup_select_num)

#Drop NAs to get model to be happy - this might be sketchy!?
y<- na.omit(x)

#Fit the full model
Full_Model <- lm(effect_f ~., data = y,
                 na.action = na.omit)
summary(Full_Model)
```

```{r, echo = FALSE}
#Stepwise model
step.model <- stepAIC(Full_Model, direction = "both", 
                      trace = FALSE)

## Remove Particle Source and Life Stage

Particle Source and Life Stage - removed from model.

#Select the columns that we want to feed into the model for simplicity
aoc_setup_select5 <- aoc_setup_select %>% 
  dplyr::select(
    `log10(dose.particles.mL.master)`, 
    `log10(dose.mg.L.master)`,
    `log10(dose.um3.mL.master)`,
    `log10(size.length.um.used.for.conversions)`, 
    # size_f,     
    poly_f, 
    shape_f,
    org_f,
    # life_f,
    bio_f,
    # `log10(exposure.duration.d)`,
    acute.chronic_f,
    # charge_f,
    # zetapotential.mV,
    # particle_source_f,
    effect_f)

#Convert all data to numerical data
aoc_setup_select_num <- data.matrix(aoc_setup_select5)

#Convert data matrix into a data frame
x <- as.data.frame(aoc_setup_select_num)

#Drop NAs to get model to be happy - this might be sketchy!?
y<- na.omit(x)

#Fit the full model
Full_Model <- lm(effect_f ~., data = y,
                 na.action = na.omit)
summary(Full_Model)
```

```{r, echo = FALSE}
#Stepwise model
step.model <- stepAIC(Full_Model, direction = "both", 
                      trace = FALSE)

## All Variables: Acute

Add all variables back in but only use acute data (fish, crustaceans, mollucs, and algae)

#Select the columns that we want to feed into the model for simplicity
aoc_setup_select5 <- aoc_setup_select %>% 
  filter(acute.chronic_f == "Acute") %>% 
  dplyr::select(
    `log10(dose.particles.mL.master)`, 
    `log10(dose.mg.L.master)`,
    `log10(dose.um3.mL.master)`,
    `log10(size.length.um.used.for.conversions)`, 
    size_f,
    poly_f, 
    shape_f,
    org_f,
    life_f,
    bio_f,
    `log10(exposure.duration.d)`,
    charge_f,
    zetapotential.mV,
    particle_source_f,
    effect_f)

#Convert all data to numerical data
aoc_setup_select_num <- data.matrix(aoc_setup_select5)

#Convert data matrix into a data frame
x <- as.data.frame(aoc_setup_select_num)

#Drop NAs to get model to be happy - this might be sketchy!?
y<- na.omit(x)

#Fit the full model
Full_Model <- lm(effect_f ~., data = y,
                 na.action = na.omit)
summary(Full_Model)
```

```{r, echo = FALSE}
#Stepwise model
step.model <- stepAIC(Full_Model, direction = "both", 
                      trace = FALSE)

## All Variables: Chronic

Add all variables back in but only use chronic data (fish, crustaceans, mollucs, and algae)

#Select the columns that we want to feed into the model for simplicity
aoc_setup_select6 <- aoc_setup_select %>% 
  filter(acute.chronic_f == "Chronic") %>% 
  dplyr::select(
    `log10(dose.particles.mL.master)`, 
    `log10(dose.mg.L.master)`,
    `log10(dose.um3.mL.master)`,
    `log10(size.length.um.used.for.conversions)`, 
    size_f,
    poly_f, 
    shape_f,
    org_f,
    life_f,
    bio_f,
    `log10(exposure.duration.d)`,
    charge_f,
    zetapotential.mV,
    particle_source_f,
    effect_f)

#Convert all data to numerical data
aoc_setup_select_num <- data.matrix(aoc_setup_select6)

#Convert data matrix into a data frame
x <- as.data.frame(aoc_setup_select_num)

#Drop NAs to get model to be happy - this might be sketchy!?
y<- na.omit(x)

#Fit the full model
Full_Model <- lm(effect_f ~., data = y,
                 na.action = na.omit)
summary(Full_Model)
```

```{r, echo = FALSE}
#Stepwise model
step.model <- stepAIC(Full_Model, direction = "both", 
                      trace = FALSE)

# Model: All Independent Variables, Response Variable: Effect (Y/N) x Effect Score

#Select the columns that we want to feed into the model for simplicity
aoc_setup_select_1 <- aoc_setup_select %>% 
  dplyr::select(
    `log10(dose.particles.mL.master)`, 
    `log10(dose.mg.L.master)`,
    `log10(dose.um3.mL.master)`,
    `log10(size.length.um.used.for.conversions)`, 
    size_f,     
    poly_f, 
    shape_f,
    org_f,
    max.size.ingest.mm,
    life_f,
    bio_f,
    `log10(exposure.duration.d)`,
    acute.chronic_f,
    charge_f,
    zetapotential.mV,
    particle_source_f,
    effect_f,
    effect.score)

#Convert all data to numerical data
aoc_setup_select_num <- data.matrix(aoc_setup_select_1)

#Convert data matrix into a data frame
x <- as.data.frame(aoc_setup_select_num)

#Drop NAs to get model to be happy - this might be sketchy!?
y<- na.omit(x)

#Fit the full model
Full_Model <- lm(effect_f*effect.score ~., data = y,
                 na.action = na.omit)
summary(Full_Model)
```

```{r, echo = FALSE}
#Stepwise model
step.model <- stepAIC(Full_Model, direction = "both", 
                      trace = FALSE)

## Remove Charge and Zeta Potential

Missing categories (NA) means that some data are incomplete. This is likely charge and zeta potential which are only in the database for PS and PC spheres. Removed from model to retain more data.

#Select the columns that we want to feed into the model for simplicity
aoc_setup_select_2 <- aoc_setup_select %>% 
  dplyr::select(
    `log10(dose.particles.mL.master)`, 
    `log10(dose.mg.L.master)`,
    `log10(dose.um3.mL.master)`,
    `log10(size.length.um.used.for.conversions)`, 
    size_f,     
    poly_f, 
    shape_f,
    org_f,
    life_f,
    bio_f,
    `log10(exposure.duration.d)`,
    acute.chronic_f,
    # charge_f,
    # zetapotential.mV,
    particle_source_f,
    effect_f,
    effect.score)

#Convert all data to numerical data
aoc_setup_select_num <- data.matrix(aoc_setup_select_2)

#Convert data matrix into a data frame
x <- as.data.frame(aoc_setup_select_num)

#Drop NAs to get model to be happy - this might be sketchy!?
y<- na.omit(x)

#Fit the full model
Full_Model <- lm(effect_f*effect.score ~., data = y,
                 na.action = na.omit)
summary(Full_Model)
```

```{r, echo = FALSE}
#Stepwise model
step.model <- stepAIC(Full_Model, direction = "both", 
                      trace = FALSE)

## Remove Size (binned)

Size (binned) not significant - removed from model.

#Select the columns that we want to feed into the model for simplicity
aoc_setup_select_3 <- aoc_setup_select %>% 
  dplyr::select(
    `log10(dose.particles.mL.master)`, 
    `log10(dose.mg.L.master)`,
    `log10(dose.um3.mL.master)`,
    `log10(size.length.um.used.for.conversions)`, 
    # size_f,     
    poly_f, 
    shape_f,
    org_f,
    life_f,
    bio_f,
    `log10(exposure.duration.d)`,
    acute.chronic_f,
    # charge_f,
    # zetapotential.mV,
    particle_source_f,
    effect_f,
    effect.score)

#Convert all data to numerical data
aoc_setup_select_num <- data.matrix(aoc_setup_select_3)

#Convert data matrix into a data frame
x <- as.data.frame(aoc_setup_select_num)

#Drop NAs to get model to be happy - this might be sketchy!?
y<- na.omit(x)

#Fit the full model
Full_Model <- lm(effect_f*effect.score ~., data = y,
                 na.action = na.omit)
summary(Full_Model)
```

```{r, echo = FALSE}
#Stepwise model
step.model <- stepAIC(Full_Model, direction = "both", 
                      trace = FALSE)

## Remove Particle Source

Particle source not significant - removed from model.

#Select the columns that we want to feed into the model for simplicity
aoc_setup_select_4 <- aoc_setup_select %>% 
  dplyr::select(
    `log10(dose.particles.mL.master)`, 
    `log10(dose.mg.L.master)`,
    `log10(dose.um3.mL.master)`,
    `log10(size.length.um.used.for.conversions)`, 
    # size_f,     
    poly_f, 
    shape_f,
    org_f,
    life_f,
    bio_f,
    `log10(exposure.duration.d)`,
    acute.chronic_f,
    # charge_f,
    # zetapotential.mV,
    # particle_source_f,
    effect_f,
    effect.score)

#Convert all data to numerical data
aoc_setup_select_num <- data.matrix(aoc_setup_select_4)

#Convert data matrix into a data frame
x <- as.data.frame(aoc_setup_select_num)

#Drop NAs to get model to be happy - this might be sketchy!?
y<- na.omit(x)

#Fit the full model
Full_Model <- lm(effect_f*effect.score ~., data = y,
                 na.action = na.omit)
summary(Full_Model)
```

```{r, echo = FALSE}
#Stepwise model
step.model <- stepAIC(Full_Model, direction = "both", 
                      trace = FALSE)

# Model: All Independent Variables + Maximum Ingestion Size, Response Variable: Effect (Y/N) x Effect Score

There is some data for the maximum ingestible size within the database. It is limited to **only 9 species**, but I wanted to try it out to see if it shows any promise.

#Select the columns that we want to feed into the model for simplicity
aoc_setup_select_1 <- aoc_setup_select %>% 
  dplyr::select(
    `log10(dose.particles.mL.master)`, 
    `log10(dose.mg.L.master)`,
    `log10(dose.um3.mL.master)`,
    `log10(size.length.um.used.for.conversions)`, 
    size_f,     
    poly_f, 
    shape_f,
    org_f,
    max.size.ingest.mm,
    life_f,
    bio_f,
    `log10(exposure.duration.d)`,
    acute.chronic_f,
    charge_f,
    zetapotential.mV,
    particle_source_f,
    effect_f,
    effect.score)

#Convert all data to numerical data
aoc_setup_select_num <- data.matrix(aoc_setup_select_1)

#Convert data matrix into a data frame
x <- as.data.frame(aoc_setup_select_num)

#Drop NAs to get model to be happy - this might be sketchy!?
y<- na.omit(x)

#Fit the full model
Full_Model <- lm(effect_f*effect.score ~., data = y,
                 na.action = na.omit)
summary(Full_Model)
```

```{r, echo = FALSE}
#Stepwise model
step.model <- stepAIC(Full_Model, direction = "both", 
                      trace = FALSE)

## Remove Charge and Zeta Potential

Missing categories (NA) means that some data are incomplete. This is likely charge and zeta potential which are only in the database for PS and PC spheres. Removed from model to retain more data.

#Select the columns that we want to feed into the model for simplicity
aoc_setup_select_2 <- aoc_setup_select %>% 
  dplyr::select(
    `log10(dose.particles.mL.master)`, 
    `log10(dose.mg.L.master)`,
    `log10(dose.um3.mL.master)`,
    `log10(size.length.um.used.for.conversions)`, 
    size_f,     
    poly_f, 
    shape_f,
    org_f,
    max.size.ingest.mm,
    life_f,
    bio_f,
    `log10(exposure.duration.d)`,
    acute.chronic_f,
    # charge_f,
    # zetapotential.mV,
    particle_source_f,
    effect_f,
    effect.score)

#Convert all data to numerical data
aoc_setup_select_num <- data.matrix(aoc_setup_select_2)

#Convert data matrix into a data frame
x <- as.data.frame(aoc_setup_select_num)

#Drop NAs to get model to be happy - this might be sketchy!?
y<- na.omit(x)

#Fit the full model
Full_Model <- lm(effect_f*effect.score ~., data = y,
                 na.action = na.omit)
summary(Full_Model)
```

```{r, echo = FALSE}
#Stepwise model
step.model <- stepAIC(Full_Model, direction = "both", 
                      trace = FALSE)

>>>>

```{r, echo = FALSE, message = FALSE, warning = FALSE}
#Packages
library(tidyverse)
library(MASS, exclude = 'select') #Stepwise regression package 
library(ggiraphExtra) #For plotting

@@

  mutate(effect_10 = case_when(
     effect_f == "Yes" ~ 1,
     effect_f == "No" ~ 0)) %>% 

@@

aoc_setup_select <- aoc_setup %>% 
  mutate(log.dose.particles.mL.master = log10(dose.particles.mL.master)) %>% 
  mutate(log.dose.mg.L.master = log10(dose.mg.L.master)) %>%
  mutate(log.dose.um3.mL.master = log10(dose.um3.mL.master)) %>%
  mutate(log.size.length.um.used.for.conversions = log10(size.length.um.used.for.conversions)) %>%
  mutate(log.exposure.duration.d = log10(exposure.duration.d))

@@

# Model: Kitchen Sink

All Independent Variables, Response Variable: Effect (Y/N)

@@

#Select the columns that we want to feed into the model for simplicity
aoc_setup_select_1 <- aoc_setup_select %>%
  filter(life_f != "Not Reported") %>% 
  filter(acute.chronic_f != "Unavailable") %>% 
  filter(size_f != "Not Reported") %>% 
  filter(poly_f != "Not Reported") %>%
  filter(shape_f != "Not Reported") %>%
  filter(environment != "Terrestrial") %>%
  dplyr::select(
    log.dose.particles.mL.master, 
    log.dose.mg.L.master,
    log.dose.um3.mL.master,
    log.size.length.um.used.for.conversions,
    # size_f,
    # poly_f,
    shape_f,
    org_f,
    life_f,
    bio_f,
    log.exposure.duration.d,
    acute.chronic_f,
    # charge_f,
    # zetapotential.mV,
    # particle_source_f,
    effect_10)

#Check for single level factors
debug_contr_error(aoc_setup_select_1)

#Omit missing data before modeling
aoc_setup_select_1 <- na.omit(aoc_setup_select_1)

#Fit the full model
Full_Model <- lm(effect_10 ~., data = aoc_setup_select_1)

summary(Full_Model)
```

@@

#Stepwise model
step.model <- stepAIC(Full_Model, direction = "both", 
                      trace = FALSE)

@@

# Model: Crustacean Fitness, Organism Level Endpoints Only

@@

#Select the columns that we want to feed into the model for simplicity
m1_crust_simple <- aoc_setup %>%
  filter(!effect_f == "NA") %>%
  filter(!acute.chronic_f == "Unavailable") %>%
  filter(bio.org == "organism") %>%
  filter(!environment == "Terrestrial") %>%
  filter(lvl1_f == "Fitness") %>%
  filter(!size_f == "Not Reported") %>%
  mutate(logdose.mg.L.master = log10(dose.mg.L.master))

#Fit the full model
m1_crust_model_dose <- glm(effect_10 ~ logdose.mg.L.master + size.length.um.used.for.conversions, data = m1_crust_simple, na.action = "na.exclude", family = "binomial")

summary(m1_crust_model_dose)
```

```{r}

#Plots
ggPredict(m1_crust_model_dose,colorn=100,jitter=FALSE)

range(m1_crust_simple$logdose.mg.L.master, na.rm = TRUE)
#[1] -11.638272   8.171552
# generate sequence of values between range in increments of 100 to give smooth appearance to model
xdose <- seq(-11.638272 ,8.171552, 0.1)
ydose <- predict(m1_crust_model_dose, 
                 list(logdose.mg.L.master = xdose), 
                 type = "response")
#ggplot with logistic data
 plot(m1_crust_simple$logdose.mg.L.master, m1_crust_simple$effect_10, pch = 16, xlab = "log dose (mg/L)", ylab = "effect")
# lines(xdose, ydose)

ggplot(m1_crust_simple, aes(x = logdose.mg.L.master, y = effect_10))+
  geom_point() +
  stat_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE)

@@
